---
title: "Asian_Final"
author: "Sayan_Patra"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## R Markdown

## Loading the data file.

```{r}
counts <- read.csv('modified_count_with_gene_name.csv')   # reading the csv file
rownames(counts) <- make.unique(counts$Symbol)   # setting the gene name as row name
counts$Symbol <- NULL                            # Removing the symbol name from the matrix as it is fixed as row name
```

## Checking the data.

```{r}
# Number of columnn in the data
ncol(counts)

# Number of features in the column
nrow(counts)
```
# Reading the metadata

```{r}
metadata<-read.csv('MODIFIED_meta_SRP301424.csv', row.names = 1)    # reading the metadata
```

```{r}
# Number of columns in the metadata
ncol(metadata)

# Number of rows in the metadata
nrow(metadata)
```

# Reading the column names in the metadata
```{r}
colnames(metadata)
```
# I am interested in the medulloblastoma.subgroup column

```{r}
metadata$medulloblastoma.subgroup
```
# I am interested in the tissue column

```{r}
metadata$tissue
```
# The medulloblastoma.subgroup represents the Normal tissues as n/a (NORM). I have to convert it into Normal only. Let's do it..............

```{r}
# I am using the gsub command for it.
metadata$medulloblastoma.subgroup <- gsub( "n/a \\(NORM\\)", "Normal", metadata$medulloblastoma.subgroup)

# Let's see
metadata$medulloblastoma.subgroup
```
# Now I am going to run the DESeq2 for Differential gene expression analysis

```{r}
dds_group <- DESeqDataSetFromMatrix(countData = counts,
                                    colData = metadata,
                                    design = ~medulloblastoma.subgroup,
                                    )
                                    smallestGroupSize <- 3  # removing the group if it has less than three column
smallestGroupSize <- 3
keep <- rowSums(counts(dds_group)>= 10) >= smallestGroupSize   #  whether the rows have 10 element and each group has 3 column
dds_group <- dds_group[keep,] # Keeping only above rows
```
# Running the DESEQ

```{r}
dds_group_01 <- DESeq(dds_group)
```
# Plotting the PCA.

# I have to ready the PCA dataset first

```{r}
# data processing for PCA
vsd <- vst(dds_group_01, blind = F)

# Plotting the PCA
pca_plot_group <- plotPCA(vsd, intgroup = c("medulloblastoma.subgroup"))

## ggplot modifications for aesthetic view
pca <- pca_plot_group+
  geom_point(aes(color = 'medulloblastoma.subgroup'),shape=21, size=3, stroke=0.5, color = 'black')
ggsave(filename = "Plots/PCA_Actual.png",
       plot = pca,
       dpi = 500)
pca
```
```{r}
# Extract results for WNT vs Normal comparison
res_subgroup1 <- results(dds_group_01, contrast = c("medulloblastoma.subgroup", "WNT", "Normal"))

# Remove rows with NA values in padj
res_filtered1 <- res_subgroup1[!is.na(res_subgroup1$padj), ]

# Filter significant genes based on adjusted p-value
genes_subgroup1 <- rownames(res_filtered1[res_filtered1$padj < 0.05, ])

# Extract results for WNT vs Normal comparison
res_subgroup2 <- results(dds_group_01, contrast = c("medulloblastoma.subgroup", "SHH", "Normal"))

# Remove rows with NA values in padj
res_filtered2 <- res_subgroup2[!is.na(res_subgroup2$padj), ]

# Filter significant genes based on adjusted p-value
genes_subgroup2 <- rownames(res_filtered2[res_filtered2$padj < 0.05, ])


# Extract results for WNT vs Normal comparison
res_subgroup3 <- results(dds_group_01, contrast = c("medulloblastoma.subgroup", "Group3", "Normal"))

# Remove rows with NA values in padj
res_filtered3 <- res_subgroup3[!is.na(res_subgroup3$padj), ]

# Filter significant genes based on adjusted p-value
genes_subgroup3 <- rownames(res_filtered3[res_filtered3$padj < 0.05, ])

# Extract results for WNT vs Normal comparison
res_subgroup4 <- results(dds_group_01, contrast = c("medulloblastoma.subgroup", "Group4", "Normal"))

# Remove rows with NA values in padj
res_filtered4 <- res_subgroup4[!is.na(res_subgroup4$padj), ]

# Filter significant genes based on adjusted p-value
genes_subgroup4 <- rownames(res_filtered4[res_filtered4$padj < 0.05, ])



gene_list <- list(
  WNT = genes_subgroup1,
  SHH = genes_subgroup2,
  Group3 = genes_subgroup3,
  Group4 = genes_subgroup4
  # Add other subgroups...
)


if (!dir.exists("Plots")) {
  dir.create("Plots")
}

png(filename = "Plots/Upset_Plot.png", width = 10, height = 8, units = "in", res = 512)

upset(fromList(gene_list),
      sets.bar.color = "blue",       # Color for the set size bars
      main.bar.color = "orange",     # Color for the intersection bars
      matrix.color = "black",      # Color for the matrix (intersection combinations)
      order.by = "freq")

dev.off()  # Close the graphics device
```


# Now Viewing results of the gene expression

```{r}
resultsNames(dds_group_01)
```
# Now generating the statistical significant genes

# for Normal vs Group 3

```{r}
Norm_vs_gr3 <- results(dds_group_01, name = "medulloblastoma.subgroup_Normal_vs_Group3")

Norm_vs_gr3
```
# Looking into the P-value of the data
```{r}
hist(Norm_vs_gr3$pvalue,
     xlab = "Normal Vs Group3 P-Value",
     ylab = "Gene Count",
     main = "Number of Gene Vs P-Value in Normal vs Group3")
```
# taking the padj = 0.05 
```{r}
Norm_vs_gr3_padj <- Norm_vs_gr3 %>% as.data.frame() %>% dplyr::filter(padj <0.05) %>% dplyr::arrange(padj)
Norm_vs_gr3_padj
```
# I am interested in upregulated and downregulated genes. Hence will set the cut off of log2FoldChange more than 1.5

# Upregulated genes

```{r}
Norm_vs_gr3_up <- Norm_vs_gr3_padj %>% filter(log2FoldChange > 1.5) %>% dplyr::arrange(desc(log2FoldChange))
Norm_vs_gr3_up
```
#Saving the upregulted in group 3 
```{r}
write.csv(Norm_vs_gr3_up, "Upregulated in Normal in Norm vs Gr3.csv", row.names = T)
```

```{r}
plotCounts(dds_group_01, gene="ZP2", intgroup = "medulloblastoma.subgroup")
```
# Downregulated Proteins

```{r}
Norm_vs_gr3_down <- Norm_vs_gr3_padj %>% filter(log2FoldChange < -1.5) %>% dplyr::arrange(log2FoldChange)
Norm_vs_gr3_down
```
# Saving the downregulated in Norm vs Group 3
```{r}
write.csv(Norm_vs_gr3_down, "Downregulated in Normal in Norm vs Gr3.csv", row.names = T)
```

# Normal vs WNT
```{r}
Norm_vs_wnt <- results(dds_group_01, contrast = c("medulloblastoma.subgroup", "Normal","WNT"))
Norm_vs_wnt
```
```{r}
Norm_vs_wnt_padj <- Norm_vs_wnt %>% as.data.frame() %>% dplyr::filter(padj <0.05) %>% dplyr::arrange(padj)

Norm_vs_wnt_padj
```
```{r}
Norm_vs_wnt_up <- Norm_vs_wnt_padj %>% filter(log2FoldChange > 1.5) %>% dplyr::arrange(desc(log2FoldChange))
Norm_vs_wnt_up
```
```{r}
write.csv(Norm_vs_wnt_up, "Upregulated in Normal in Normal vs WNT.csv", row.names = T)
```

# Downregulated in Normal vs WNT
```{r}
Norm_vs_wnt_down <- Norm_vs_wnt_padj %>% filter(log2FoldChange < -1.5) %>% dplyr::arrange(log2FoldChange)
Norm_vs_wnt_down
```
```{r}
write.csv(Norm_vs_wnt_down, "Downregulated in Normal in Normal vs wnt.csv", row.names = T)
```

```{r}
Norm_vs_shh <- results(dds_group_01, contrast = c("medulloblastoma.subgroup", "Normal","SHH"))
Norm_vs_shh
```
```{r}
Norm_vs_shh_padj <- Norm_vs_shh %>% as.data.frame() %>% filter(padj < 0.05) %>% dplyr::arrange(padj)
Norm_vs_shh_padj
```
```{r}
Norm_vs_shh_up <- Norm_vs_shh_padj %>% filter(log2FoldChange > 1.5) %>% dplyr::arrange(desc(log2FoldChange))
Norm_vs_shh_up
```
```{r}
write.csv(Norm_vs_shh_up, "Upregulated in Normal in Norm vs SHH.csv", row.names = T)
```

```{r}
Norm_vs_shh_down <- Norm_vs_shh_padj %>% filter(log2FoldChange < -1.5) %>% dplyr::arrange((log2FoldChange))
Norm_vs_shh_down
```
```{r}
write.csv(Norm_vs_shh_down,"Downregulated in Normal in Norm vs Shh.csv", row.names = T)
```

```{r}
Norm_vs_gr4 <- results(dds_group_01, contrast = c("medulloblastoma.subgroup", "Normal", "Group4"))
Norm_vs_gr4
```
```{r}
Norm_vs_gr4_padj <- Norm_vs_gr4 %>% as.data.frame() %>% filter(padj < 0.05) %>% dplyr::arrange(padj)
Norm_vs_gr4_padj
```
```{r}
Norm_vs_gr4_up <- Norm_vs_gr4_padj %>% filter(log2FoldChange > 1.5) %>% dplyr::arrange(desc(log2FoldChange))
Norm_vs_gr4_up
```
```{r}
Norm_vs_gr4_down <- Norm_vs_gr4_padj %>% filter(log2FoldChange < -1.5) %>% dplyr::arrange(log2FoldChange)
Norm_vs_gr4_down
```
```{r}
#up
write.csv(Norm_vs_gr4_up, "Upregulated in Normal in Normal vs Group4.csv", row.names = T)
#down
write.csv(Norm_vs_gr4_down,"Downregulated in Normal in Normal vs Group 4.csv", row.names = T)
```

```{r}
wnt_vs_shh <- results(dds_group_01, contrast = c("medulloblastoma.subgroup", "WNT", "SHH"))
wnt_vs_shh
```
```{r}
wnt_vs_shh_padj <- wnt_vs_shh %>% as.data.frame() %>% filter(padj < 0.05) %>% dplyr::arrange(padj)
wnt_vs_shh_padj
```
```{r}
# upregualte in wnt
wnt_vs_shh_up <- wnt_vs_shh_padj %>% filter(log2FoldChange > 1.5) %>% dplyr::arrange(desc(log2FoldChange))
wnt_vs_shh_up

# downregulated in wnt
wnt_vs_shh_down <- wnt_vs_shh_padj %>% filter(log2FoldChange < -1.5) %>% dplyr::arrange(log2FoldChange)
wnt_vs_shh_down

```
```{r}
# up
write.csv(wnt_vs_shh_up, "Upregulated in WNT in wnt vs shh.csv", row.names = T)

# down
write.csv(wnt_vs_shh_down,"Downregulated in wnt in wnt vs shh.csv", row.names = T)
```

```{r}
wnt_vs_gr3 <- results(dds_group_01, contrast = c("medulloblastoma.subgroup", "WNT", "Group3"))
wnt_vs_gr3

wnt_vs_gr3_padj <- wnt_vs_gr3 %>% as.data.frame() %>% filter(padj < 0.05) %>% dplyr::arrange(padj)
wnt_vs_gr3_padj

```
```{r}
# up
wnt_vs_gr3_up <- wnt_vs_gr3_padj %>% filter(log2FoldChange > 1.5) %>% dplyr::arrange(desc(log2FoldChange))
wnt_vs_gr3_up

# down
wnt_vs_gr3_down <- wnt_vs_gr3_padj %>% filter(log2FoldChange < -1.5) %>% dplyr::arrange(log2FoldChange)
wnt_vs_gr3_down
```
```{r}
# up
write.csv(wnt_vs_gr3_up, "Upregulated in WNT in wnt vs group3.csv", row.names = T)

#down
write.csv(wnt_vs_gr3_down, "Downregulated in WNT in wnt vs group3.csv", row.names = T)
```

```{r}
wnt_vs_gr4 <- results(dds_group_01, contrast = c("medulloblastoma.subgroup", "WNT", "Group4"))
wnt_vs_gr4

wnt_vs_gr4_padj <- wnt_vs_gr4 %>% as.data.frame() %>% filter(padj < 0.05) %>% dplyr::arrange(padj)
wnt_vs_gr4_padj
```
```{r}
# up
wnt_vs_gr4_up <- wnt_vs_gr4_padj %>% filter(log2FoldChange > 1.5) %>% dplyr::arrange(desc(log2FoldChange))
wnt_vs_gr4_up

# down
wnt_vs_gr4_down <- wnt_vs_gr4_padj %>% filter(log2FoldChange < -1.5) %>% dplyr::arrange(log2FoldChange)
wnt_vs_gr4_down
```
```{r}
# up
write.csv(wnt_vs_gr4_up, "Upregulated in WNT in wnt vs group4.csv", row.names = T)

#down
write.csv(wnt_vs_gr4_down, "Downregulated in WNT in wnt vs group4.csv", row.names = T)
```


```{r}
plotCounts(dds_group_01, gene="LINC00879", intgroup = "medulloblastoma.subgroup")
```
```{r}
shh_vs_gr3 <- results(dds_group_01, contrast = c("medulloblastoma.subgroup", "SHH", "Group3"))
shh_vs_gr3

shh_vs_gr3_padj <- shh_vs_gr3 %>% as.data.frame() %>% filter(padj < 0.05) %>% dplyr::arrange(padj)
shh_vs_gr3_padj
```

```{r}
# up
shh_vs_gr3_up <- shh_vs_gr3_padj %>% filter(log2FoldChange > 1.5) %>% dplyr::arrange(desc(log2FoldChange))
shh_vs_gr3_up

# down
shh_vs_gr3_down <- shh_vs_gr3_padj %>% filter(log2FoldChange < -1.5) %>% dplyr::arrange(log2FoldChange)
shh_vs_gr3_down
```
```{r}
# up
write.csv(shh_vs_gr3_up, "Upregulated in SHH in Shh vs Group3.csv", row.names = T)

#down
write.csv(shh_vs_gr3_down,"Downregulated in SHH in SHH vs Group3.csv", row.names = T)
```

```{r}
plotCounts(dds_group_01, gene="LOC105374142", intgroup = "medulloblastoma.subgroup")
```
```{r}
shh_vs_gr4 <- results(dds_group_01, contrast = c("medulloblastoma.subgroup", "SHH", "Group4"))
shh_vs_gr4

shh_vs_gr4_padj <- shh_vs_gr4 %>% as.data.frame() %>% filter(padj < 0.05) %>% dplyr::arrange(padj)
shh_vs_gr4_padj
```
```{r}
# up
shh_vs_gr4_up <- shh_vs_gr4_padj %>% filter(log2FoldChange > 1.5) %>% dplyr::arrange(desc(log2FoldChange))
shh_vs_gr4_up

# down
shh_vs_gr4_down <- shh_vs_gr4_padj %>% filter(log2FoldChange < -1.5) %>% dplyr::arrange(log2FoldChange)
shh_vs_gr4_down
```
```{r}
# up
write.csv(shh_vs_gr4_up, "Upregulated in SHH in Shh vs Group4.csv", row.names = T)

#down
write.csv(shh_vs_gr4_down,"Downregulated in SHH in SHH vs Group4.csv", row.names = T)
```

```{r}
plotCounts(dds_group_01, gene="LOC105374142", intgroup = "medulloblastoma.subgroup")
```
```{r}
gr3_vs_gr4 <- results(dds_group_01, contrast = c("medulloblastoma.subgroup", "Group3", "Group4"))
gr3_vs_gr4

gr3_vs_gr4_padj <- gr3_vs_gr4 %>% as.data.frame() %>% filter(padj < 0.05) %>% dplyr::arrange(padj)
gr3_vs_gr4_padj
```
```{r}
# up
gr3_vs_gr4_up <- gr3_vs_gr4_padj %>% filter(log2FoldChange > 1.5) %>% dplyr::arrange(desc(log2FoldChange))
gr3_vs_gr4_up

# down
gr3_vs_gr4_down <- gr3_vs_gr4_padj %>% filter(log2FoldChange < -1.5) %>% dplyr::arrange(log2FoldChange)
gr3_vs_gr4_down
```
```{r}
# up
write.csv(gr3_vs_gr4_up, "Upregulated in Group3 in Group3 vs Group4.csv", row.names = T)

#down
write.csv(gr3_vs_gr4_down,"Downregulated in Group3 in Group3 vs Group4.csv", row.names = T)
```

```{r}
plotCounts(dds_group_01, gene="LINC00261", intgroup = "medulloblastoma.subgroup")
```

# Correlation Matrix and Plot
```{r, height = 100, width = 100}
Normalized <- counts(dds_group_01, normalized = T)

# Correlation Matrix
cor_matrix <- cor(Normalized, method = "pearson")

# Correlation Plot
png(filename="Plots/03_Correlation_Plot.png", width=5000, height=5000, res=500)
pheatmap(cor_matrix, clustering_distance_rows="correlation", clustering_distance_cols="correlation", height = 800, width = 800)
dev.off()
```



























